---
title: "additional-analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{additional-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
```

These pieces of analysis code either require some level of manual curation,
require multiple data sets to be loaded at once, are specific to a species, or
otherwise are not suited for the more automated analysis code provided in the
sample-analysis vignette.

# Regional Prevalence of blaOXA-23 family alleles in *A. baumannii*
```{r, eval = FALSE}
ab %>%
  filter(oxa_family == "blaOXA-23") %>%
  count_isolates(region) %>%
  rename(oxa23fam_containing = isolates) %>%
  left_join(ab %>% count_isolates(region), by = "region") %>%
  mutate(proportion = oxa23fam_containing / isolates) %>%
  readr::write_tsv(file = file.path(manual_output_path, "ab_blaOXA23_regional.tsv"))
```

# Regional prevalence of (assumed) acquired blaOXA allels in *P. aeruginosa*
```{r, eval = FALSE}
pa %>%
  tidyr::drop_na(oxa_family) %>%
  filter(oxa_family != "blaOXA-50") %>%
  count_isolates(region) %>%
  rename(acquired_oxa = isolates) %>%
  left_join(pa %>% count_isolates(region), by = "region") %>%
  mutate(proportion = acquired_oxa / isolates) %>%
  readr::write_tsv(file = file.path(manual_output_path, "pa_acquired_blaOXA_regional.tsv"))
```

# Carbapenemases in *A. baumannii* excluding blaOXA-51 family alleles
```{r, eval = FALSE}

# Allele Counts
carbapenemase_allele_counts <- ab %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  count_isolates(allele)

carbapenemase_allele_counts %>%
  readr::write_csv(file.path(manual_output_path, "ab_carbapenemase_allele_counts_without_blaOXA-51_family.csv"))

# Gene Counts (Without blaOXA-51 Family)
carbapenemase_gene_counts <- ab %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  count_isolates(gene)

carbapenemase_gene_counts %>%
  readr::write_csv(file.path(manual_output_path, "ab_carbapenemase_gene_counts_without_blaOXA-51_family.csv"))

# Presence by Region
carbapenemase_presence_by_region <- ab %>%
  group_by(region, region_markdown) %>%
  mutate(isolates = n_distinct(biosample)) %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  summarize(
    isolates = mean(isolates),
    carbapenemase_alleles = length(allele[subclass == "CARBAPENEM"]),
    carbapenemase_isolates = n_distinct(biosample[subclass == "CARBAPENEM"]),
    carbapenemase_isolate_proportion = carbapenemase_isolates / isolates
  )

carbapenemase_presence_by_region %>%
  dplyr::select(-region_markdown) %>%
  readr::write_csv(file.path(manual_output_path, "ab_carbapenemase_presence_by_region_without_blaOXA-51_family.csv"))

ab_carbapenemase_presence_by_region_without_blaOXA51_family_plot <- carbapenemase_presence_by_region %>%
  mutate(non_carbapenemase_isolates = isolates - carbapenemase_isolates) %>%
  dplyr::rename(Present = carbapenemase_isolates, Absent = non_carbapenemase_isolates) %>%
  tidyr::pivot_longer(cols = c(Present, Absent), names_to = "carbapenemase") %>%
  tidyr::drop_na(region) %>%
  group_by(region, region_markdown) %>%
  top_n_by_group(carbapenemase, count = value, n = 3, relevel = "name") %>%
  donut_plot_panel(fill = carbapenemase, counts = value, facet = region_markdown,
                   legend_title = "**Carbapenemase**", count_label = "Isolates", nrow = 2,
                   legend.position = "bottom")

  ggplot2::ggsave(filename = file.path(
    manual_output_path,
    paste0("ab_carbapenemase_presence_by_region_without_blaOXA-51_family.svg")),
    plot = ab_carbapenemase_presence_by_region_without_blaOXA51_family_plot
  )

# Presence Over Time
carbapenemase_presence_over_time <- ab %>%
  group_by(period) %>%
  mutate(isolates = n_distinct(biosample)) %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  summarize(
    isolates = mean(isolates),
    carbapenemase_alleles = length(allele[subclass == "CARBAPENEM"]),
    carbapenemase_isolates = n_distinct(biosample[subclass == "CARBAPENEM"]),
    carbapenemase_isolate_prop = carbapenemase_isolates / isolates
  )

carbapenemase_presence_over_time %>%
  readr::write_csv(file.path(manual_output_path, "ab_carbapenemase_presence_over_time_without_blaOXA-51_family.csv"))


# Presence Over Time (Since 2000)
carbapenemase_presence_over_time_since_2000 <- ab %>%
  group_by(period_2000) %>%
  mutate(isolates = n_distinct(biosample)) %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  summarize(
    isolates = mean(isolates),
    carbapenemase_alleles = length(allele[subclass == "CARBAPENEM"]),
    carbapenemase_isolates = n_distinct(biosample[subclass == "CARBAPENEM"]),
    carbapenemase_isolate_prop = carbapenemase_isolates / isolates
  )

carbapenemase_presence_over_time_since_2000 %>%
  readr::write_csv(file.path(manual_output_path, "ab_carbapenemase_presence_over_time_since_2000_without_blaOXA-51_family.csv"))

```

# ADCs of Particular Interest
```{r, eval = FALSE}
# blaADC-223 Time Increase MLST
ab %>%
  tidyr::drop_na(period_2000) %>%
  dplyr::filter(allele == "blaADC-223") %>%
  count_isolates(mlst, period_2000)

# blaADC-182 MLST
ab %>%
  tidyr::drop_na(period_2000) %>%
  dplyr::filter(allele == "blaADC-182") %>%
  count_isolates(mlst)

# blaADC-183 MLST
ab %>%
  tidyr::drop_na(period_2000) %>%
  dplyr::filter(allele == "blaADC-183") %>%
  count_isolates(mlst)

# blaADC-235 MLST
ab %>%
  tidyr::drop_na(period_2000) %>%
  dplyr::filter(allele == "blaADC-235") %>%
  count_isolates(mlst)
```


# Combined count of isolates by country
```{r, eval = FALSE}
combined_country_isolate_counts <- ab %>%
  count_isolates(standardized_country) %>%
  rename(ab = isolates) %>%
  full_join(pa %>% count_isolates(standardized_country), by = "standardized_country") %>%
  rename(pa = isolates) %>%
  tidyr::replace_na(list(ab = 0L, pa = 0L)) %>%
  mutate(total = ab + pa) %>%
  arrange(desc(total)) %>%
  mutate(standardized_country = forcats::fct_reorder(standardized_country, total, I)) %>%
  arrange(desc(standardized_country))

combined_country_isolate_counts %>%
  readr::write_csv(file.path(manual_output_path, "combined_country_isolate_counts.csv"))

combined_country_isolate_counts_50plus <- combined_country_isolate_counts %>%
  filter(pa >= 50 | ab >= 50) %>%
  tidyr::drop_na(standardized_country)

combined_country_isolate_counts_50plus %>%
  readr::write_csv(file.path(manual_output_path, "combined_country_isolate_counts_50plus.csv"))
```

# Combined count of isolates by region
```{r, eval = FALSE}
combined_region_isolate_counts <- ab %>%
  count_isolates(region) %>%
  rename(ab = isolates) %>%
  full_join(pa %>% count_isolates(region), by = "region") %>%
  rename(pa = isolates) %>%
  tidyr::replace_na(list(ab = 0L, pa = 0L)) %>%
  mutate(total = ab + pa) %>%
  arrange(desc(total)) %>%
  mutate(region = forcats::fct_reorder(region, total, I)) %>%
  arrange(desc(region))

combined_region_isolate_counts %>%
  readr::write_csv(file.path(manual_output_path, "combined_region_isolate_counts.csv"))

```

# Comparison to POP Study
```{r, eval = FALSE}
# Portion of isolates
pa %>%
  summarize(
    pop_2018 = n_distinct(biosample[year == 2018 & bioproject == "PRJNA824880"]),
    pop_2019 = n_distinct(biosample[year == 2019 & bioproject == "PRJNA824880"]),
    pop_total = n_distinct(biosample[bioproject == "PRJNA824880" & year %in% 2018:2019]),
    nonpop_2018 = n_distinct(biosample[year == 2018 & bioproject != "PRJNA824880"]),
    nonpop_2019 = n_distinct(biosample[year == 2019 & bioproject != "PRJNA824880"]),
    nonpop_total = n_distinct(biosample[bioproject != "PRJNA824880" & year %in% 2018:2019]),
    overall_2018 = n_distinct(biosample[year == 2018]),
    overall_2019 = n_distinct(biosample[year == 2019]),
    overall_total = n_distinct(biosample[year %in% 2018:2019])
  )

# Overall presence of carbapenemase genes
pa %>%
  filter(year == 2018 | year == 2019) %>%
  filter(bioproject != "PRJNA824880") %>%
  summarize(
    isolates = n_distinct(biosample),
    carbapenemase_alleles = length(allele[subclass == "CARBAPENEM"]),
    carbapenemase_isolates = n_distinct(biosample[subclass == "CARBAPENEM"]),
    carbapenemase_isolate_proportion = carbapenemase_isolates / isolates
  )

# Overall presence of carbapenemase genes
pa %>%
  filter(year == 2018 | year == 2019) %>%
  filter(bioproject != "PRJNA824880") %>%
  group_by(region) %>%
  summarize(
    isolates = n_distinct(biosample),
    carbapenemase_alleles = length(allele[subclass == "CARBAPENEM"]),
    carbapenemase_isolates = n_distinct(biosample[subclass == "CARBAPENEM"]),
    carbapenemase_isolate_proportion = carbapenemase_isolates / isolates
  )

# Regional presence of carbapenemase genes
pa %>%
  filter(year == 2018 | year == 2019) %>%
  filter(bioproject != "PRJNA824880") %>%
  group_by(region) %>%
  summarize(
    isolates = n_distinct(biosample),
    carbapenemase_alleles = length(allele[subclass == "CARBAPENEM"]),
    carbapenemase_isolates = n_distinct(biosample[subclass == "CARBAPENEM"]),
    carbapenemase_isolate_proportion = carbapenemase_isolates / isolates
  )

# Counts by carbapenemase gene
pa %>%
  filter(year == 2018 | year == 2019) %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(bioproject != "PRJNA824880") %>%
  distinct(biosample, gene) %>%
  count(gene, name = "isolates") %>%
  mutate(overall_proportion = isolates / length(unique(pa %>% filter(year == 2018 | year == 2019) %>% pull(biosample)))) %>%
  mutate(carpabenemase_proportion = isolates / length(unique(pa %>% filter(year == 2018 | year == 2019) %>% filter(subclass == "CARBAPENEM") %>% pull(biosample)))) %>%
  arrange(desc(isolates))

# Counts by sequence type
pa %>%
  filter(year == 2018 | year == 2019) %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(bioproject != "PRJNA824880") %>%
  count_isolates(mlst)
```
# Comparison to TEST Study
```{r, eval = FALSE}
# Overall in data set
ab %>%
  filter(bioproject == "PRJEB27899") %>%
  count_isolates(bioproject)

# Isolates counts (Note that years are not provided in the BioSample metadata for the TEST isolates)
ab %>%
  summarize(
    test_total = n_distinct(biosample[bioproject == "PRJEB27899"]),
    nontest_total_2012_2016 = n_distinct(biosample[bioproject != "PRJEB27899" & year %in% 2012:2016]),
    test_prop = test_total / nontest_total_2012_2016
  )

# Presence of blaOXA-23 and blaOXA-40 families

ab %>%
  filter(bioproject != "PRJEB27899" & year %in% 2012:2016) %>%
  count_isolates(oxa_family) %>%
  mutate(prop = isolates / (ab %>% filter(bioproject != "PRJEB27899" & year %in% 2012:2016) %>% distinct(biosample) %>% nrow()))

# Presence of blaNDM-1 and blaIMP-26
ab %>%
  filter(bioproject != "PRJEB27899" & year %in% 2012:2016) %>%
  filter(allele %in% c("blaNDM-1", "blaIMP-26")) %>%
  count_isolates(allele) %>%
  mutate(prop = isolates / (ab %>% filter(bioproject != "PRJEB27899" & year %in% 2012:2016) %>% distinct(biosample) %>% nrow()))

# Novel balOXA-51 Family Alleles
test_biosamples <- ab %>%
  filter(bioproject != "PRJEB27899" & year %in% 2012:2016) %>%
  distinct(biosample) %>%
  pull(biosample)

ab %>%
  filter(biosample %in% test_biosamples) %>%
  filter(oxa_family == "blaOXA-51") %>%
  filter_assigned_bla %>%
  distinct(ipg) %>%
  nrow()

# Count of ST
ab %>%
  filter(biosample %in% test_biosamples) %>%
  count_isolates(mlst) %>%
  mutate(prop = isolates / sum(isolates)) %>%
  arrange(desc(isolates))
```

# Comparison to SNAP Study
```{r, eval = FALSE}
# Isolates counts
c(
    overall_2017 = ab %>% filter(year == 2017) %>% distinct(biosample) %>% nrow(),
    overall_2018 = ab %>% filter(year == 2018) %>% distinct(biosample) %>% nrow(),
    overall_2019 = ab %>% filter(year == 2019) %>% distinct(biosample) %>% nrow(),
    overall_total = ab %>% filter(year %in% 2017:2019) %>% distinct(biosample) %>% nrow(),
    snap_2017 = ab %>% filter(year == 2017) %>% filter (bioproject == "PRJNA906166") %>% distinct(biosample) %>% nrow(),
    snap_2018 = ab %>% filter(year == 2018) %>% filter (bioproject == "PRJNA906166") %>% distinct(biosample) %>% nrow(),
    snap_2019 = ab %>% filter(year == 2019) %>% filter (bioproject == "PRJNA906166") %>% distinct(biosample) %>% nrow(),
    snap_total = ab %>% filter(year %in% 2017:2019) %>% filter (bioproject == "PRJNA906166") %>% distinct(biosample) %>% nrow(),
    nonsnap_2017 = ab %>% filter(year == 2017) %>% distinct(biosample) %>% nrow() - ab %>% filter(year == 2017) %>% filter (bioproject == "PRJNA906166") %>% distinct(biosample) %>% nrow(),
    nonsnap_2017 = ab %>% filter(year == 2018) %>% distinct(biosample) %>% nrow() - ab %>% filter(year == 2018) %>% filter (bioproject == "PRJNA906166") %>% distinct(biosample) %>% nrow(),
    nonsnap_2017 = ab %>% filter(year == 2019) %>% distinct(biosample) %>% nrow() - ab %>% filter(year == 2019) %>% filter (bioproject == "PRJNA906166") %>% distinct(biosample) %>% nrow(),
    nonsnap_total = ab %>% filter(year %in% 2017:2019) %>% distinct(biosample) %>% nrow() - ab %>% filter(year %in% 2017:2019) %>% filter (bioproject == "PRJNA906166") %>% distinct(biosample) %>% nrow()
  )

# SNAP Biosamples
snap_biosamples <- ab %>%
  filter(bioproject == "PRJNA906166") %>%
  distinct(biosample) %>%
  pull(biosample)

#Acquired carbapenemase allele-containing isolate count
ab %>%
  filter(year %in% 2017:2019) %>%
  filter(!(biosample %in% snap_biosamples)) %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  distinct(biosample, .keep_all = TRUE) %>%
  nrow()

#Acquired carbapenemase OXA family counts
ab %>%
  filter(year %in% 2017:2019) %>%
  filter(!(biosample %in% snap_biosamples)) %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  count(oxa_family)

#Acquired carbapenemase gene counts
ab %>%
  filter(year %in% 2017:2019) %>%
  filter(!(biosample %in% snap_biosamples)) %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  count(gene, allele)


#blaOXA-23 allele-containing isolate count
ab %>%
  filter(year %in% 2017:2019) %>%
  filter(!(biosample %in% snap_biosamples)) %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  filter(allele == "blaOXA-23") %>%
  distinct(biosample, .keep_all = TRUE) %>%
  nrow()

#blaOXA-24 allele-containing isolate coun
ab %>%
  filter(year %in% 2017:2019) %>%
  filter(!(biosample %in% snap_biosamples)) %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  filter(allele == "blaOXA-24") %>%
  distinct(biosample, .keep_all = TRUE) %>%
  nrow()

#blaOXA-24 family-containing isolate coun
ab %>%
  filter(year %in% 2017:2019) %>%
  filter(!(biosample %in% snap_biosamples)) %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  filter(oxa_family == "blaOXA-24") %>%
  distinct(biosample, .keep_all = TRUE) %>%
  nrow()

# Sequence types in Overall
ab %>%
  filter(!(biosample %in% snap_biosamples)) %>%
  distinct(biosample, .keep_all = TRUE) %>%
  filter(year %in% 2017:2019) %>%
  count(mlst) %>%
  mutate(proportion = n / sum(n)) %>%
  arrange(desc(n))

# Sequence types in USA
ab %>%
  filter(!(biosample %in% snap_biosamples)) %>%
    distinct(biosample, .keep_all = TRUE) %>%
  filter(year %in% 2017:2019) %>%
  filter(location_broad == "USA") %>%
  count(mlst) %>%
  mutate(proportion = n / sum(n)) %>%
  arrange(desc(n))

# Sequence types in South America
ab %>%
  filter(!(biosample %in% snap_biosamples)) %>%
  distinct(biosample, .keep_all = TRUE) %>%
  filter(year %in% 2017:2019) %>%
  filter(region == "South America") %>%
  count(mlst) %>%
  mutate(proportion = n / sum(n)) %>%
  arrange(desc(n))


```
# Generate list of alleles to download for sequence comparisons
```{r, eval = FALSE}
# Load Reference Gene Catalog file
rgc <- readr::read_delim(file.path(data_path, "refgene.tsv"), delim = "\t", escape_double = FALSE, trim_ws = TRUE, show_col_types = FALSE)

# ADC
adc_rgc <- rgc %>%
  dplyr::select(allele, gene_family, refseq_protein_accession) %>%
  dplyr::filter(gene_family == "blaADC") %>%
  dplyr::pull(refseq_protein_accession)

adc_mbe <- ab %>%
  filter_unassigned_bla() %>%
  dplyr::filter(grepl("blaADC", allele)) %>%
  distinct(ipg_accession, .keep_all = TRUE) %>%
  dplyr::pull(protein)

readr::write_lines(na.omit(c(adc_rgc, adc_mbe)),
                   file = file.path(manual_output_path, "protein_accessions_adc.txt"))
  
# OXA-51 Family
oxa_51_rgc <- rgc %>%
  dplyr::mutate(oxa_family = dplyr::case_when(
    grepl('family', product_name) ~ stringr::str_extract(product_name, "OXA-\\d+?(?= family)"),
    TRUE ~ stringr::str_extract(product_name, "OXA-\\d+")
  )) %>%
  dplyr::select(allele, oxa_family, refseq_protein_accession) %>%
  dplyr::filter(oxa_family == "OXA-51") %>%
  dplyr::pull(refseq_protein_accession)

oxa_51_mbe <- ab %>%
  filter_unassigned_bla() %>%
  dplyr::filter(oxa_family == "blaOXA-51") %>%
  distinct(ipg_accession, .keep_all = TRUE) %>%
  dplyr::pull(protein)

readr::write_lines(na.omit(c(oxa_51_rgc, oxa_51_mbe)),
                   file = file.path(manual_output_path, "protein_accessions_oxa-51.txt"))

# PDC
pdc_rgc <- rgc %>%
  dplyr::select(allele, gene_family, refseq_protein_accession) %>%
  dplyr::filter(gene_family == "blaPDC") %>%
  dplyr::pull(refseq_protein_accession)

pdc_mbe <- pa %>%
  filter_unassigned_bla() %>%
  dplyr::filter(grepl("blaPDC", allele)) %>%
  distinct(ipg_accession, .keep_all = TRUE) %>%
  dplyr::pull(protein)

readr::write_lines(na.omit(c(pdc_rgc, pdc_mbe)),
                   file = file.path(manual_output_path, "protein_accessions_pdc.txt"))

# OXA-50 Family
oxa_50_rgc <- rgc %>%
  dplyr::mutate(oxa_family = dplyr::case_when(
    grepl('family', product_name) ~ stringr::str_extract(product_name, "OXA-\\d+?(?= family)"),
    TRUE ~ stringr::str_extract(product_name, "OXA-\\d+")
  )) %>%
  dplyr::select(allele, oxa_family, refseq_protein_accession) %>%
  dplyr::filter(oxa_family == "OXA-50") %>%
  dplyr::pull(refseq_protein_accession)

oxa_50_mbe <- pa %>%
  filter_unassigned_bla() %>%
  dplyr::filter(oxa_family == "blaOXA-50") %>%
  distinct(ipg_accession, .keep_all = TRUE) %>%
  dplyr::pull(protein)

readr::write_lines(na.omit(c(oxa_50_rgc, oxa_50_mbe)),
                   file = file.path(manual_output_path, "protein_accessions_oxa-50.txt"))
```

# ADC-33 with OXA-23
```{r, eval = FALSE}
adc33_biosamples <- ab %>%
  filter_assigned_bla() %>%
  dplyr::filter(allele == "blaADC-33") %>%
  dplyr::pull(biosample)
  
adc33_oxa23 <- ab %>%
  filter_assigned_bla() %>%
  dplyr::filter(biosample %in% adc33_biosamples) %>%
  dplyr::filter(oxa_family == "blaOXA-23") %>%
  dplyr::distinct(biosample) %>%
  nrow()

adc33_oxa23_text <- c(
  paste0("Total blaADC-33: ",length(adc33_biosamples)),
  paste0("With blaOXA-23 Family Members: ", adc33_oxa23),
  paste0("Proportion: ",adc33_oxa23 / length(adc33_biosamples)) 
)

readr::write_lines(adc33_oxa23_text,
                   file = file.path(manual_output_path, "adc-33_oxa-23.txt"))

```
