---
title: "additional-analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{additional-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
```

These pieces of analysis code either require some level of manual curation,
require multiple data sets to be loaded at once, are specific to a species, or
otherwise are not suited for the more automated analysis code provided in the
sample-analysis vignette.

# Regional Prevalence of blaOXA-23 family alleles in *A. baumannii*
```{r, eval = FALSE}
ab %>%
  filter(oxa_family == "blaOXA-23") %>%
  count_isolates(region) %>%
  rename(oxa23fam_containing = isolates) %>%
  left_join(ab %>% count_isolates(region), by = "region") %>%
  mutate(proportion = oxa23fam_containing / isolates) %>%
  readr::write_tsv(file = file.path(manual_output_path, "ab_blaOXA23_regional.tsv"))
```

# Regional prevalence of (assumed) acquired blaOXA alleels in *P. aeruginosa*
```{r, eval = FALSE}
pa %>%
  tidyr::drop_na(oxa_family) %>%
  filter(oxa_family != "blaOXA-50") %>%
  count_isolates(region) %>%
  rename(acquired_oxa = isolates) %>%
  left_join(isolates %>% count_isolates(region), by = "region") %>%
  mutate(proportion = acquired_oxa / isolates) %>%
  readr::write_tsv(file = file.path(manual_output_path, "pa_acquired_blaOXA_regional.tsv"))
```

# Carbapenemases in *A. baumannii* excluding blaOXA-51 family alleles
```{r, eval = FALSE}

# Allele Counts
carbapenemase_allele_counts <- ab %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  count_isolates(allele)

carbapenemase_allele_counts %>%
  readr::write_csv(file.path(manual_output_path, "ab_carbapenemase_allele_counts_without_blaOXA-51_family.csv"))

# Gene Counts (Without blaOXA-51 Family)
carbapenemase_gene_counts <- ab %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  count_isolates(gene)

carbapenemase_gene_counts %>%
  readr::write_csv(file.path(manual_output_path, "ab_carbapenemase_gene_counts_without_blaOXA-51_family.csv"))

# Presence by Region
carbapenemase_presence_by_region <- ab %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  distinct(biosample, .keep_all = TRUE) %>%
  count_isolates(region, region_markdown) %>%
  rename(carbapenemase_isolates = isolates) %>%
  left_join(ab %>% count_isolates(region), by = "region") %>%
  rename(total_isolates = isolates) %>%
  mutate(proportion = carbapenemase_isolates / total_isolates)

carbapenemase_presence_by_region %>%
  dplyr::select(-region_markdown) %>%
  readr::write_csv(file.path(manual_output_path, "ab_carbapenemase_presence_by_region_without_blaOXA-51_family.csv"))

carbapenemase_presence_by_region %>%
  mutate(non_carbapenemase_isolates = total_isolates - carbapenemase_isolates) %>%
  dplyr::rename(Present = carbapenemase_isolates, Absent = non_carbapenemase_isolates) %>%
  tidyr::pivot_longer(cols = c(Present, Absent), names_to = "carbapenemase") %>%
  tidyr::drop_na(region) %>%
  group_by(region, region_markdown) %>%
  top_n_by_group(carbapenemase, count = value, n = 3, relevel = "name") %>%
  donut_plot_panel(fill = carbapenemase, counts = value, facet = region_markdown,
                   legend_title = "**Carbapenemase**", count_label = "Isolates", nrow = 2,
                   legend.position = "bottom")

# Presence Over Time
carbapenemase_presence_over_time <- ab %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  count_isolates(period) %>%
  rename(carbapenemase_isolates = isolates) %>%
  left_join(ab %>% count_isolates(period), by = "period") %>%
  rename(total_isolates = isolates) %>%
  mutate(proportion = carbapenemase_isolates / total_isolates) %>%
  relevel_numeric(period) %>%
  arrange(period)

carbapenemase_presence_over_time %>%
  readr::write_csv(file.path(manual_output_path, "ab_carbapenemase_presence_over_time_without_blaOXA-51_family.csv"))


# Presence Over Time (Since 2000)
carbapenemase_presence_over_time_since_2000 <- ab %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  count_isolates(period_2000) %>%
  rename(carbapenemase_isolates = isolates) %>%
  left_join(ab %>% count_isolates(period_2000), by = "period_2000") %>%
  rename(total_isolates = isolates) %>%
  mutate(proportion = carbapenemase_isolates / total_isolates) %>%
  relevel_numeric(period_2000) %>%
  arrange(period_2000)

carbapenemase_presence_over_time_since_2000 %>%
  readr::write_csv(file.path(manual_output_path, "ab_carbapenemase_presence_over_time_since_2000_without_blaOXA-51_family.csv"))

# Presence by Region
carbapenemase_presence_by_region <- pa %>%
  filter(subclass == "CARBAPENEM") %>%
  filter(!grepl("blaOXA-51", oxa_family)) %>%
  distinct(biosample, .keep_all = TRUE) %>%
  count_isolates(region, region_markdown) %>%
  rename(carbapenemase_isolates = isolates) %>%
  left_join(pa %>% count_isolates(region), by = "region") %>%
  rename(total_isolates = isolates) %>%
  mutate(proportion = carbapenemase_isolates / total_isolates)

carbapenemase_presence_by_region_plot <- carbapenemase_presence_by_region %>%
  mutate(non_carbapenemase_isolates = total_isolates - carbapenemase_isolates) %>%
  dplyr::rename(Present = carbapenemase_isolates, Absent = non_carbapenemase_isolates) %>%
  tidyr::pivot_longer(cols = c(Present, Absent), names_to = "carbapenemase") %>%
  tidyr::drop_na(region) %>%
  group_by(region, region_markdown) %>%
  top_n_by_group(carbapenemase, count = value, n = 3, relevel = "name") %>%
  donut_plot_panel(fill = carbapenemase, counts = value, facet = region_markdown,
                   legend_title = "**Carbapenemase**", count_label = "Isolates", nrow = 2,
                   legend.position = "bottom")

ggplot2::ggsave(filename = file.path(
  manual_output_path,
  paste0("ab_carbapenemase_presence_by_region_no_blaOXA51-family.svg")),
  plot = carbapenemase_presence_by_region_plot
)


```

# Combined count of isolates by country
```{r, eval = FALSE}
combined_country_isolate_counts <- pa %>%
  count_isolates(standardized_country) %>%
  rename(pa = isolates) %>%
  full_join(ab %>% count_isolates(standardized_country), by = "standardized_country") %>%
  rename(ab = isolates)

combined_country_isolate_counts %>%
  readr::write_csv(file.path(manual_output_path, "combined_country_isolate_counts.csv"))
```

# Combined count of isolates by country (only countries with >= 50 isolates for one species)
```{r, eval = FALSE}
combined_country_isolate_counts_50plus <- pa %>%
  count_isolates(standardized_country) %>%
  rename(pa = isolates) %>%
  full_join(ab %>% count_isolates(standardized_country), by = "standardized_country") %>%
  rename(ab = isolates) %>%
  filter(pa >= 50 | ab >= 50)

combined_country_isolate_counts_50plus %>%
  readr::write_csv(file.path(manual_output_path, "combined_country_isolate_counts_50plus.csv"))
```

