---
title: "Determining Sequence Types"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mlst-determination}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
```

## Introduction
Sequence types should be determiend using FastMLST, which is available from
https://github.com/EnzoAndree/FastMLST and described in "FastMLST: A Multi-core
Tool for Multilocus Sequence Typing of Draft Genome Assemblies" by
Guerrero-Araya et al. available at
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8637782/.

The output from FastMLST can be easilly reprocessed using the following
procedure to provide more granular detail about indodidual isolates and to
update sequence types to newer versions of the database without needing to
re-run FastMLST.

## Running FastMLST
See the official focumentation for FastMLST at
https://github.com/EnzoAndree/FastMLST/blob/master/README.md for guidence on how
to run FastMLST and the various options available.

The only specific requirment for use with the `pdallele` package is that the
output must be in CSV format.

## Reprocessing FastMLST Output
```{r, eval = FALSE}

mlst <- reprocess_mlst("/Users/armack/Library/CloudStorage/OneDrive-CaseWesternReserveUniversity/Allele Distribution Data/Pseudomonas/2023-11-15/pa_mlst.csv")
```
